---
title: "Assignment_4"
author: "Rishitha Reddy Muddasani"
date: "2023-03-18"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

Loading the data set and the Libraries
```{r}
library(flexclust)
library(cluster)
library(tidyverse)
library(factoextra)
library(FactoMineR)
library(tinytex)
library(ggcorrplot)

P_Data<-read.csv("~/Downloads/Pharmaceuticals.csv")
P_Data<-na.omit(P_Data)
```
*TASK 1*
The 21 firms are grouped using the numerical variables (1–9).
```{r}
row.names(P_Data)<-P_Data[,1]
Clustering_dataset<-P_Data[,3:11]
```
data scalability
```{r}
set.seed(143)
Scaled_data<-scale(Clustering_dataset)
```

Kmeans computation using random K values
```{r}
set.seed(143)
kmeans_2_centers<-kmeans(Scaled_data,centers = 2, nstart = 15)
kmeans_4_centers<-kmeans(Scaled_data,centers = 4, nstart = 15)
kmeans_8_centers<-kmeans(Scaled_data,centers = 8, nstart = 15)
plot_kmeans_2_centers<-fviz_cluster(kmeans_2_centers,data = Scaled_data) + ggtitle("K=2")
plot_kmeans_4_centers<-fviz_cluster(kmeans_4_centers,data = Scaled_data) + ggtitle("K=4")
plot_kmeans_8_centers<-fviz_cluster(kmeans_8_centers,data = Scaled_data) + ggtitle("K=8")
plot_kmeans_2_centers
plot_kmeans_4_centers
plot_kmeans_8_centers
```

Finding the optimal K appropriate for clustering using WSS and Silhouette

```{r}
wss<-fviz_nbclust(Scaled_data,kmeans,method="wss")
silhouette<-fviz_nbclust(Scaled_data,kmeans,method="silhouette")
wss
silhouette
distance<-dist(Scaled_data,metho='euclidean')
fviz_dist(distance)
```
k is 2 from WSS and 5 from silhouette. The number 5 ensures that the sum of squires inside each cluster is minimal and that there is considerable spacing between them.

*TASK 2*

Using Kmeans to find an appropriate k 
```{r}
set.seed(143)
kmeans_5_centers<-kmeans(Scaled_data,centers = 5, nstart = 10)
kmeans_5_centers
plot_kmeans_5_centers<-fviz_cluster(kmeans_5_centers,data = Scaled_data) + ggtitle("K=5")
plot_kmeans_5_centers
```


```{r}
Clustering_dataset_1<-Clustering_dataset%>% mutate(Cluster_no=kmeans_5_centers$cluster)%>% group_by(Cluster_no)%>%summarise_all('mean')
Clustering_dataset_1
```


Following clusters have been created for companies:

Cluster_1= ABT,AHM,AZN,BMY,LLY,NVS,SGP,WYE

Cluster_2= BAY,CHTT,IVX

Cluster_3=AVE,ELN,MRX,WPI

Cluster_4=AGN,PHA

Cluster_5=GSK,JNJ,MRK,PFE


This can be inferred from the clusters that were generated.

1. Cluster 1 contains a collection of businesses with a modest return on equity and return on investment.

2. Cluster 2 Companies have extremely low ROA, ROE, market capitalization, and asset turnover. This means that these businesses are exceedingly dangerous.

3. Similar to cluster 2, Cluster 3 features group corporations, but with slightly lower risk.
 
4. Companies in cluster 4 are more risky than those in cluster 2 because they have very good PE ratios but weak ROA and ROE.

5. Companies in Cluster 5 have excellent ROE, ROA, and market capitalization.


*TASK 3*

```{r}
#Is there a pattern in the clusters with respect to the numerical 
#variables (10 to 12)? (those \n #not used in forming the clusters)
Clustering_datase_2<- P_Data[,12:14] %>% mutate(Clusters=kmeans_5_centers$cluster)
ggplot(Clustering_datase_2, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters',y='Frequency')
ggplot(Clustering_datase_2, mapping = aes(factor(Clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='Clusters',y='Frequency')
ggplot(Clustering_datase_2, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters',y='Frequency')
```

Clusters and the variable Median Recommendation exhibit a pattern, as can be observed. Similar to what the second cluster shows between moderate buy and hold, the third cluster recommends between moderate purchase and moderate sell. The majority of pharmaceutical businesses are based in the US, as can be seen from the location graph, although there isn't much of a pattern there. With the exception of the bulk of companies being listed on NYSE, there is no discernible relationship between clusters and exchanges.


*TASK 4* - Naming clusters:

[It is done based on the net Market capitalization(size) and Return on Assets(money)]

Cluster 1: Large-Thousands
Cluster 2: Extra Small-Penny
Cluster 3: Small- Dollars
Cluster 4: Medium-Hundreds
Cluster 5: Extra Large-Millions 
